---
title: "CuttingGardens 2023 Carbon footprint analysis"
output: html_notebook
---

Working with results of [this form](https://docs.google.com/forms/d/1dGbXsicczUR3OZ2Hd-3asfE7MmpwIMUqrtk5AqOVxew/edit)

## Read data
```{r READ DATA, message=FALSE}
library(tidyverse)
library(ggmap)
library(geosphere)
library(patchwork)
register_google(key=read_file('google_key.txt'))
googlesheets4::gs4_auth(email = 'maximilien.chaumon@gmail.com')


# dum <- geocode('Paris, Fr')
deaccent <- function(c){
  c %>%
    str_replace('é','e')%>%
    str_replace('ü','ue')%>%
    str_replace('ö','oe') %>%
    str_replace('ä','ae') %>%
    str_replace('á','a') %>%
    str_replace('italia','italy')
}

Gardens <- googlesheets4::read_sheet(ss = "https://docs.google.com/spreadsheets/d/1KnPOV7WEIkpPIXy8CLGDY_sIdVI-wC77VXtwqiMN9RM/edit#gid=1863576514",sheet = "Official Gardens' list!",range = "A1:P22" ) %>% 
  select(`Garden City`, Country, `Number of participants`, lon, lat) %>% 
  rename(garden.city = `Garden City`, garden.country = `Country`, N = `Number of participants`) %>%
  mutate(across(is.character, trimws),
         across(is.character, str_to_lower),
         across(is.character,deaccent)) %>%
  unite(garden, garden.city, garden.country, sep = ' ; ', remove = F)
  
d <- readr::read_csv('Réponses_Gardens2023.csv',show_col_types = F) %>%
  rename(garden.city = 2,
         origin = 3,
         transport_in = 4,
         transport_out = 5) %>%
  filter(garden.city != "I only attended online") %>%
  filter( ! if_any(contains('transport'), ~ str_detect(.x,'online'))) %>%
  filter( ! if_any(contains('transport'), ~  str_detect(.x,'Other'))) %>%
  select(-1,-6) %>%
  separate(col = origin, into = c('origin.city','origin.country'), sep=';', remove=F) %>%
  mutate(across(is.character, trimws),
         across(is.character, str_to_lower),
         across(is.character,deaccent),
         transport_out = ifelse(transport_out == 'same as i came!', transport_in, transport_out),
         across(starts_with('transport'), ~ str_replace(.x, 'city bus / intercity coach', 'city bus')),
         across(starts_with('transport'), ~ str_replace(.x, 'regional train', 'train'))) %>%
  unite(col = origin, origin.city, origin.country, sep=' ; ', remove=F)  %>%
  mutate(origin = ifelse(is.na(origin), garden.city, origin)) %>%
  left_join(Gardens %>% select(garden, garden.city, garden.country), by = "garden.city")


location_codes <- read_csv('locationCodes.csv',col_types = 'ccdd') %>%
  mutate(across(is.character, str_to_lower)) %>%
  mutate(across(is.character, trimws)) %>%
  bind_rows(Gardens %>% rename(origin = garden, city = garden.city, country = garden.country) %>% select(-N)) %>%
  unite(col = origin, city, country, sep=' ; ', remove=F)  %>%
  distinct(origin,.keep_all = TRUE) %>%
  filter(!(is.na(city) | is.na(country)))

allcities <- levels(as.factor(d$origin))

for (c in allcities) {
  c <- deaccent(c)
  if (! c %in% location_codes$origin) {
    lonlat <- geocode(c)
    cc <- str_split(c, ' ; ')
    city <- cc[[1]][1]
    country <- cc[[1]][2]
    location_codes <- add_row(location_codes,
                              origin = c,
                              city = city,
                              country = country,
                              lon = lonlat$lon,
                              lat = lonlat$lat)
    write_csv(location_codes %>%
                arrange(country, city) %>%
                select(country, city, lon, lat),'locationCodes.csv')
  }
}


country_codes <- read_csv('countryCodes.csv',col_types = 'ddc')  %>%
  distinct(.keep_all = TRUE)
allcountries <- levels(as.factor(d$origin.country))

for (c in allcountries) {
  if (! c %in% country_codes$country) {
    lonlat <- geocode(c)
    country_codes <- add_row(country_codes,
                              country = c,
                              lon = lonlat$lon,
                              lat = lonlat$lat)
    write_csv(country_codes,'countryCodes.csv')
  }
}
location_codes <- location_codes %>%
  rename(origin.lon = lon, origin.lat = lat) %>%
  arrange(country, city) %>%
  select(-city, -country)
Gardens <- Gardens %>%
  rename(garden.lon = lon, garden.lat = lat) %>%
  select(-garden.city, -garden.country)
country_codes <- country_codes %>%
  rename(country.lon = lon, country.lat = lat)


d <- d %>% left_join(location_codes,by='origin') %>%
  left_join(country_codes, by = join_by(origin.country == country)) %>%
  left_join(Gardens %>% select(-N), by = 'garden') %>%
  mutate(origin.garden.distance_km = distCosine(as.matrix(cbind(origin.lon,origin.lat)),
                          as.matrix(cbind(garden.lon, garden.lat))) / 1000)
         

```

## Response counts per garden

```{r count responses, rows.print=100}
library(kableExtra)

tokable <- d %>% mutate(Garden = str_to_title(garden), .keep = 'none') %>%
  group_by(Garden) %>%
  summarize(nResponses = n()) %>%
  left_join(Gardens %>%
              mutate(Garden = str_to_title(garden)) %>%
              select(Garden,N) %>%
              rename(NParticipants = N),
            by = join_by(Garden)) %>%
  bind_rows(tibble(Garden = 'Total', nResponses = sum(.$nResponses), NParticipants = sum(.$NParticipants))) %>%
  mutate(proportionResponses = round(nResponses / NParticipants * 100 ))%>%
  select(Garden, NParticipants, nResponses, proportionResponses)

tokable %>%
  kable(caption = 'Number of responses per Garden, and proportion of participants who responded', col.names = c('Garden','Participants','Responses','Proportion (%)')) %>%
  kable_styling(latex_options = "striped") %>%
  row_spec(nrow(tokable), bold = T)

```

## Analysis

For each responding garden, we estimate the amount of CO2 emitted by transportation. Following this, we estimate what it would have been if all participants had been travelling to a single location.

### Read Emission Factors

```{r}
EF <- readr::read_csv("CO2PerPersonTransport.csv", col_types = 'dc') %>%
  mutate(across(is.character, str_to_lower)) %>%
  mutate(across(is.character, trimws)) 
EF %>%
  mutate(transport = str_replace(transport,'Metro Or Rer', 'Metro or Regional Train')) %>%
  mutate(transport = str_to_title(transport)) %>%
  select(transport, EF) %>%
  kable(caption = 'Emission factors for all recorded means of transportation.', col.names = c('Transportation means', 'Emission factor (kg / km)'),
        digits = 3) %>%
  kable_styling(latex_options = "striped")

```

### Compute emissions of reported transportation without any resampling

```{r, rows.print=100}

nud <- d %>% 
  left_join(EF, by = join_by(transport_in == transport)) %>%
  rename(EF_in = EF) %>% 
  left_join(EF, by = join_by(transport_out == transport)) %>%
  rename(EF_out = EF) %>%
  rename(Garden = garden) %>%
  mutate(CO2_in = EF_in * origin.garden.distance_km,
         CO2_out = EF_out * origin.garden.distance_km,
         CO2_tot = CO2_in + CO2_out)

nud %>% group_by(Garden) %>%
  mutate(Garden = str_to_title(Garden)) %>%
  summarize(CO2_tot = sum(CO2_tot)) %>%
  bind_rows(tibble(Garden = 'Total', CO2_tot = sum(.$CO2_tot))) %>%
  kable(caption = 'Total CO2 emissions per Garden, in kg', col.names = c('Garden','CO2 emissions (kg)'),
        digits = 0) %>%
  kable_styling(latex_options = "striped")
```

## Estimation tenant compte des effectifs

Dans chaque Garden, on resample parmi les moyens de transport utilisés

```{r, rows.print=100}
nrep <- 100
# On reprend les facteurs d'émission
nud <- d %>% 
  left_join(EF, by = join_by(transport_in == transport)) %>%
  rename(EF_in = EF) %>% 
  left_join(EF, by = join_by(transport_out == transport)) %>%
  rename(EF_out = EF) %>%
  mutate(CO2_in = EF_in * origin.garden.distance_km,
         CO2_out = EF_out * origin.garden.distance_km,
         CO2_tot = CO2_in + CO2_out) %>%
  left_join(Gardens, by = 'garden')

# for each garden, bootstrap resample the right number of pple, compute CO2, store
rsmpd <- tibble()
# for each garden
for (g in levels(as.factor(nud$garden))) {
  # focus on data of that garden
  tmp <- filter(nud,garden == g)
  if (sum(tmp$CO2_tot) == 0) next
  # get the number of participants
  N <- Gardens %>%
    filter(garden == g) %>%
    .$N
  # for each resampling
  for (isim in 1:nrep) {
    # pick with replacement N participants
    rsmpd <- tmp[sample(1:nrow(tmp),size = N, replace = T),] %>%
      summarize(CO2_tot = sum(CO2_tot),
                garden = first(garden),
                garden.city = first(garden.city),
                garden.country = first(garden.country)) %>%
      bind_rows(rsmpd)
  }
}
# at this point, we have nrep resampled CO2 emission estimates for each garden
# now compute the mean and sd of these estimates for each garden
rsmpds <- rsmpd %>%
  group_by(garden, garden.city, garden.country) %>%
  summarize(sdCO2_tot = sd(CO2_tot),
            mCO2_tot = mean(CO2_tot))
rsmpds %>% mutate(mCO2_tot = round(mCO2_tot),
              sdCO2_tot = round(sdCO2_tot)) %>%
  ungroup()%>%
  select(garden, mCO2_tot, sdCO2_tot) %>%
  bind_rows(tibble(garden = 'Total',
                   mCO2_tot = sum(.$mCO2_tot),
                   sdCO2_tot = round(sqrt(sum(.$sdCO2_tot^2)))))
```


```{r, rows.print=100}
rsmpds %>%
  ungroup() %>%
  select(garden, mCO2_tot, sdCO2_tot) %>%
  bind_rows(tibble(garden = 'Total',
                   mCO2_tot = sum(.$mCO2_tot),
                   sdCO2_tot = round(sqrt(sum(.$sdCO2_tot^2))))) %>%
  kable(caption = 'Estimated CO2eq emissions due to transportation in the Cutting Gardens Scenario', col.names = c('Garden', 'Average CO2 emissions (kg)', 'standard deviation (kg)'),
        digits = 0) %>%
  kable_styling(latex_options = "striped") %>%
  row_spec(1:nrow(rsmpds)+1, extra_css = "padding: 0px") %>%
  row_spec(nrow(rsmpds)+1, bold = T)

# a barplot based on rsmpds with x = garden, y = emissions, color = garden, with errorbars = sd
ggplot(rsmpds %>% mutate(mCO2_tot = mCO2_tot / 1000, sdCO2_tot = sdCO2_tot / 1000), aes(x=garden, y=mCO2_tot, fill=garden)) + 
  geom_bar(stat="identity", width=1, color = 'black', show.legend = F) +
  geom_errorbar(aes(ymin=mCO2_tot-sdCO2_tot, ymax=mCO2_tot+sdCO2_tot), width=.2,
                position=position_dodge(1)) +
  coord_flip() +
  labs(x = NULL, y = expression(paste('CO2 emissions (10'^'3',' kg)')), fill = NULL) +
  theme_minimal()

# # plot rsmpd with ggplot boxplot
# ggplot(rsmpd, aes(x="", y=CO2_tot, fill=garden)) +
#   geom_violin() + facet_wrap(~garden, scales = "free_y")

# ggplot(rsmpds, aes(x="", y=mCO2_tot, fill=garden)) +
#   geom_bar(stat="identity", width=1) +
#   coord_polar("y", start=0)  +
#   geom_text(aes(label = paste(garden.country, round(mCO2_tot))), position = position_stack(vjust=0.5)) +
#   labs(x = NULL, y = NULL, fill = NULL)

# ggplot(rsmpd,aes(x=CO2_tot,fill = garden)) +
#   geom_histogram(position='dodge', bins = 10, show.legend = F) +
#   facet_wrap(~garden,
#              scales = "free_x")

```

## Now assuming a single central location in Paris

```{r}
TheGarden <- tibble(garden.city = 'Paris', garden.country = 'France') %>%
  mutate(across(is.character, str_to_lower)) %>%
  mutate(across(is.character, trimws)) %>% 
  unite(garden, garden.city, garden.country, sep = ' ; ', remove = F) %>%
  left_join(location_codes, by = join_by( garden == origin)) %>%
  rename(garden.lon = origin.lon, garden.lat = origin.lat)

```


Assuming all \< a certain distance use train, above use plane

```{r}
alldmin <- seq(300,1500,10)
nrep <- 10

toplot = tibble()
for (dmin in alldmin) { 
  nud <- d %>%
    rename_with(.cols = starts_with('garden'),.fn =  ~ paste0('actual.',.x)) %>%
    bind_cols(TheGarden) %>%
    mutate(origin.garden.distance_km = distCosine(as.matrix(cbind(origin.lon,origin.lat)),
                                                  as.matrix(cbind(garden.lon, garden.lat))) / 1000) %>%
    select(-starts_with('garden')) %>%
    rename_with(.cols = starts_with('actual'),.fn =  ~ str_remove(.x,'actual.')) %>%
    mutate(transport_in = ifelse(origin.garden.distance_km < dmin, 'train','plane'),
           transport_out = ifelse(origin.garden.distance_km < dmin, 'train','plane')) %>%
    left_join(EF, by = join_by(transport_in == transport)) %>%
    rename(EF_in = EF) %>% 
    left_join(EF, by = join_by(transport_out == transport)) %>%
    rename(EF_out = EF) %>%
    mutate(CO2_in = EF_in * origin.garden.distance_km,
           CO2_out = EF_out * origin.garden.distance_km,
           CO2_tot = CO2_in + CO2_out)
  
  
  # for each garden, bootstrap resample the right number of pple, compute CO2, store
  rsmpd <- tibble()
  # for each garden
  for (g in levels(as.factor(nud$garden))) {
    # focus on data of that garden
    tmp <- filter(nud,garden == g)
    if (sum(tmp$CO2_tot) == 0) next
    # get the number of participants
    N <- Gardens %>%
      filter(garden == g) %>%
      .$N
    # for each resampling
    for (isim in 1:nrep) {
      # pick with replacement N participants
      rsmpd <- tmp[sample(1:nrow(tmp),size = N, replace = T),] %>%
        summarize(CO2_tot = sum(CO2_tot),
                  garden = first(garden),
                  garden.city = first(garden.city),
                  garden.country = first(garden.country)) %>%
        bind_rows(rsmpd)
    }
  }
  # at this point, we have nrep resampled CO2 emission estimates for each garden
  # now compute the mean and sd of these estimates for each garden
  rsmpds <- rsmpd %>%
    group_by(garden, garden.city, garden.country) %>%
    summarize(sdCO2_tot = sd(CO2_tot),
              mCO2_tot = mean(CO2_tot), .groups = 'drop')
  # rsmpds %>% mutate(mCO2_tot = round(mCO2_tot),
  #                   sdCO2_tot = round(sdCO2_tot)) %>%
  #   ungroup()%>%
  #   select(garden, mCO2_tot, sdCO2_tot) %>%
  #   bind_rows(tibble(garden = 'Total',
  #                    mCO2_tot = sum(.$mCO2_tot),
  #                    sdCO2_tot = round(sqrt(sum(.$sdCO2_tot^2)))))
  
  
  tmp <- nud %>% group_by(garden) %>%
    summarize(CO2_tot = round(sum(CO2_tot))) %>%
    bind_rows(tibble(garden = 'Total',
                     CO2_tot = sum(.$CO2_tot)))
  
  toplot <- tmp %>% filter(garden == 'Total') %>%
    mutate(dmin = dmin) %>%
    select(dmin,garden,CO2_tot) %>%
    bind_rows(toplot)
}

p1 <- ggplot(toplot, aes(x=dmin, y=CO2_tot / 1000)) +
  geom_line() +
  # geom_point() +
  # scale_x_reverse() +
  labs(x = 'Distance threshold for flying (km)', y = 'Total CO2 emissions (tons)', title = 'Conference CO2 emissions as a function of \ndistance for switching from train to plane') + 
  theme_minimal()

p2 <- ggplot(nud, aes(x = origin.garden.distance_km, fill = origin.country)) +
  geom_histogram(bins = 30) +
  labs(x = 'Distance to Paris (km)', y = 'frequency', title = 'Distance to Paris') +
  theme_minimal() #+ theme(legend.position = 'bottom', legend.box="vertical", legend.margin=margin())

p1
p2

```



